% ****************************************************************************************************
% Business Model Concept NEW
% ****************************************************************************************************
\begin{longtable}{L{\column}L{\column}L{\column}L{\column}}
	
	\toprule
	\endfirsthead
	\toprule
	\multicolumn{4}{l}{\textit{continued from previous page (\nameref{bm:concept})}}\\ 
	\midrule
	\endhead
	\midrule
	\multicolumn{4}{r}{\textit{continued on next page}}\\
	\bottomrule
	\endfoot
	\bottomrule
	\caption[Business Model Conceptualization]{Business Model Conceptualization adapted from \citet[p. 54]{Johnson2008}}
	\label{bm:concept}
	\endlastfoot
	
	\multicolumn{4}{c}{\textbf{Customer Value Proposition (CVP)}}\\ \midrule
	
	\textbf{\small Target customer} &
	\small \textbf{Job to be done} to solve an important problem or fulfill an important need for the target customer. &
	\multicolumn{2}{L{\columnT}}{\small \textbf{Offering}, which satisfies the problem or fulfills the need. This is defined not only by what is sold but also by how it's sold.} \\ 

% KEY RESOURCES
\multicolumn{2}{L{\columnT}}{\textbf{Key Resources} \small needed to deliver the customer value proposition profitably. Might include:
\begin{itemize}[leftmargin=*, parsep=0pt, topsep=0pt, itemsep=0pt]
	\item \textbf{People}
	\item \textbf{Technology, products}
	\item \textbf{Equipment}
	\item \textbf{Information}
	\item \textbf{Channels}
	\item \textbf{Partnerships, alliances}
	\item \textbf{Brand}\vspace{-\baselineskip} 
\end{itemize}
} &
\multicolumn{2}{L{\columnT}}{\textbf{Key Processes} \small as well as rules, metrics, and norms, that make the profitable delivery of the customer value proposition repeatable and scalable. Might include:
\begin{itemize}[leftmargin=*, parsep=0pt, topsep=0pt, itemsep=0pt]
	\item \textbf{Processes:} design, product development, sourcing, manufacturing, marketing, hiring and training, IT
	\item \textbf{Rules and metrics:} margin requirements for investment, credit terms, lead times, supplier terms
	\item \textbf{Norms:} opportunity size needed for investment, approach to customers and channels\vspace{-\baselineskip} 
\end{itemize}} \\ \midrule

% PROFIT FORMULA
\multicolumn{4}{L{\columnF}}{\textbf{Profit Formula} \small
\begin{itemize}[leftmargin=*, parsep=0pt, topsep=0pt, itemsep=0pt]
		\item \textbf{Revenue model:} How much money can be made: price x volume. Volume can be thought of in terms of market size, purchase frequency, ancillary sales, etc.
		\item \textbf{Cost structure:} How costs are allocated: includes cost of key assets, direct costs, indirect costs, economies of scale.
		\item \textbf{Margin model:} How much each transaction should net to achieve desired profit levels. 
		\item \textbf{Resource velocity:} How quickly resources need to be used to support target volume. Includes lead times, throughput, inventory turns, asset utilization, and so on.\vspace{-\baselineskip} 
\end{itemize} 
}\\ 
\end{longtable}



